Distributional Trees for Circular Data

### 2019-01-29 pkg/disttree/inst/IWSM2019/abstract_IWSM2019.tex
further revisions
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@2412 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: e3783c19a54ce64b2e08630bc05e6b1f16cc9a91

\section{Motivation}
Circular response variables occur in a variety of applications and subject areas. 
For example, gun crime data on a $24$-hour scale is analysed in the social-economics, 
animal orientation or gene-structure analysis are often subject of examination in biology, 
and wind direction is one of the most important weather variables in meteorology.

One way to represent such a two-dimensional direction is by considering angles in a
unity circle. This requires specifying a starting point (angle 0) and defining
either a clockwise or anti-clockwise rotation. However, these settings can vary
and hence should not influence statistical inference from a fitted model.
For that reason circular data demands for purpose-built statistical methods. 
One common way of modelling circular data is to apply the von Mises distribution 
which is often regarded as the circular analog of the normal distribution. 
Estimating its two distribution parameter by linear functions of regressors is
a possible approach, however, might not be able to detect abrupt effects.
As an alternative  we propose the application of distributional trees to
model wind directions based on a set of regression variables.

>>>>

\section{Motivation}
Circular response variables occur in a variety of applications and subject areas. 
For example, gun crime data on a $24$-hour scale is analysed in the social-economics, 
animal orientation or gene-structure analysis are often subject of examination in biology, 
and wind direction is one of the most important weather variables in meteorology.
Representing such a two-dimensional direction by an angle in a unity circle requires 
the specification of a starting point (angle 0) and either a clockwise or anti-clockwise 
rotation. However, these settings can vary but should not influence statistical 
inference from a fitted model.
For that reason circular data demands for purpose-built statistical methods. 
One common way of modelling circular data is to apply the von Mises distribution 
which is often regarded as the circular analog of the normal distribution. 
Estimating its two distribution parameter by linear functions of regressor variables 
is a possible approach, however, might not be able to detect abrupt effects.
As an alternative  we propose the application of distributional trees to
model wind directions based on a set of regression variables.
